Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
Standardized citations

Antibodies are highly specialized proteins of the immune system that
are produced in response to a foreign substance, called an antigen. A
mature antibody binds a specific antigen with high affinity, while
only weakly interacting with other antigens, or not at all. This high
affinity, high specificity sets it apart from other pharmaceuticals.
Furthermore, in contrast with small drug molecules, antibodies can not
only bind pockets, but also flat, concave or even convex surfaces
\cite{Tavers2001}. Their unique characteristics have enabled
researchers to develop efficient antibody drugs for treating cancers,
autoimmune disorders, infectious diseases and many more \cite{Lu2020}.
Their ability to target an immense variety of antigens allows for
endless possibilities in application. The global market size was
valued at USD 130.9 billion in 2020, estimated to grow 223.7 billion
by the end of 2025 at a compound annual growth rate of 11.31\%
\cite{market2020}. Four of the top 10 best-selling drugs in 2020 were
monoclonal antibodies \cite{Urquhart2021}.

In order to rationally design therapeutic antibodies, knowledge of
their structure is essential. The acquired structural information can
be used to increase binding affinity to a target of interest,
predicting both the exact binding site and the antibody stability as
well as assessing immunogenicity \cite{Abhinandan2007}. As
experimental structure determination is very costly and time
consuming, computational predictions of an antibody's structure are
used to streamline the process.

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly achievable
\cite{North2011},\cite{Weitzner2015}. However, the CDR loop 3 of the
heavy chain (CDR-H3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable \cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops \cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-Angstrom non-antibody
structural neighbour, as well as that 30\% of CDR-H3 loops have a
completely unique structure compared with under 3\% for all other
loops on average \cite{Regep2017}.

Apart from being the most structurally diverse, the CDR-H3 loop is also
the most important for antigen binding, being located at the center of
the binding site and forming the most contacts to the antigen
\cite{MacCallum1996}. In fact, it was demonstrated that differences in
this loop alone were sufficient to enable otherwise identical
antibodies to distinguish between various antigens \cite{Xu2000}.

@@

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDR-H3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDR-H3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above. \protect\cite{Wong2015}}  
  \label{fig:loopdist}
\end{figure}

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDR-H3 loops. This was also shown by the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build structural models from antibody sequences. The CDR-H3 loop
modelling quality achieved at the contests was on average much lower
for loops of longer lengths \cite{Almagro2011},\cite{Almagro2014}.

Several different approaches for generating three-dimensional
structure models from antibody sequences exist such as RosettaAntibody
\cite{Sircar2009},\cite{Sivasubramanian2009},
ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017} and abYmod,
UCL's in-house software developed by Prof. Martin. One of the most
used methods is RosettaAntibody, which implements template selection
and ab initio CDR-H3 loop modelling using loop fragments and employing
specific angle restraints which bias the conformational space towards
so-called ‘kinked' loops\cite{Schoeder2021},\cite{Weitzner2017}. In
contrast, ABodyBuilder uses a database search algorithm
(FREAD\cite{Choi2010}) for CDR loop modelling.

In this project abYmod is used, which is to be found at
http://abymod.abysis.org/. The program utilizes extensive canonical
class definitions, VH/VL angle prediction and a large database of loop
structures (LoopDB) for CDR-H3 modelling to achieve optimal
results. Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues \cite{Martin1996}and where
an exact match is not possible, a nearest match is made. Then, the
program identifies the 10 best overall-matching PDB files according to
sequence identity for the light and heavy chain. Of these, the best
template is then identified for each CDR using sequence similarity and
identity. In general, modelling the antibody using the single best
overall-matching template works best and the CDR-specific templates
are used if there is no canonical match. The VH/VL packing angle is
then determined either using machine learning or the chosen template
structure for one of the chains. CDR specific templates, if selected,
are grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using the LoopDB database,
containing CDR-H3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results or outperforms other modelling software
(see results section).

@@

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.CDR-H2) of the antibody
model. The given score is the probability that a specific region
(e.g. CDR-H2) will be will be modeled within a specific RMSD threshold
\cite{Leem2016}. Thus, it can be used to obtain an expected RMSD
value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data.\cite{Leem2016} ABLooper also provides a
confidence metric for the CDR-H3 loop model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}.However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality model. Furthermore, it remains unclear
how well the generated diversity score reflects model
quality\cite{Abanades2022}.

@@

\section{Results}
To asses the predictive power regarding the CDR-H3 loop, the software
was tested on a test-set of antibody structures used in the 2014 and
2011 Antibody Modelling Assessments
\cite{Almagro2014},\cite{Almagro2011}. As the results depicted in
\ref{fig:AMA} show, abYmod achieves results similar to, or better than
other modelling programs. However, the outliers with very high RMSD
values increase abYmod's RMSD average. The predictor in this work
would aim to identify such outlier models.

The predictive power of any machine learning model is largely
dependent on the quality and size of the dataset it was trained on. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{Almagro2011},
established by Prof. Andrew Martin, was utilised (1924 non-redundant
structures). The root-mean-square deviation (RMSD) value, a measure of
distance between backbone C-$\alpha$ atoms of superimposed crystal structures
and modelled structures, is calculated (see methods). This metric for
model quality was used to classify models.

@@

The full pipeline for creating the final machine learning model that
will predict model quality by giving its RMSD range starts with a
feature-set calculation using the antibody sequence. The feature set
includes many different attributes linked to sequence, physical
characteristics, interactions, etc. within as well as outside of the
loop. There is a plethora of information to train our classifier on,
including packing quality, protrusion, hydrophobicity, pseudo-energy
(see protrusion, hydrophobicity, pseudo-energy (see methods for
details). The sequence logo \ref{fig:logo} visualizes amino acid
occurrence within the loop sequence, elements of which can be
extracted as features \cite{Thomsen2012},\cite{Shaner1993}.

@@

A validation set, usually used for testing during the optimization
stage will be omitted in favour of stratified K-folds cross-validation
(k=10) \cite{Krstajic2014} \cite{Kohavi1995}.

\subsection{Model Assessment} 
Model assessment must be considered at two levels as performance
metrics of binary and multi-class classifiers are calculated
differently and must thus be considered separately. The Mathews
Correlation Coefficient (MCC) \cite{Matthews1975} is deemed the most
informative, taking the ratios of the four confusion matrix categories
into account 39and is thus more reliable than the F1 score and
accuracy . It is also consistent for both binary and multi-class
problems and therefore well suited for our purpose.

@@

\begin{figure}
  \includegraphics[scale=0.6]{angle.eps}
  \caption {Diagram visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e H95 and H102, shown as
    red spheres) of the CDR-H3 (shown in red) are identified. Then, a
    line is drawn between the two C$\alpha$ atoms of these residues (yellow
    dashed line). The distance of the C$\alpha$-atom of each residue in the
    CDR-H3 loop to this line is calculated. The residue which has the
    greatest distance to the line (shown as green sphere) is outputted
    as one-letter amino acid code and used as feature. The distance in
    \AA\ (depicted as yellow arrow) is used as the ‘protrusion'
    feature. }
  \label{fig:angle}
\end{figure}

>>>>

Antibodies are highly specialized proteins of the immune system that
are produced in response to a foreign substance, called an antigen. A
mature antibody binds a specific antigen with high affinity, while
only weakly interacting with other antigens, or not at all. This high
affinity, high specificity sets it apart from other pharmaceuticals.
Furthermore, in contrast with small drug molecules, antibodies can not
only bind pockets, but also flat, concave or even convex surfaces\cite{Tavers2001}. Their unique characteristics have enabled
researchers to develop efficient antibody drugs for treating cancers,
autoimmune disorders, infectious diseases and many more\cite{Lu2020}.
Their ability to target an immense variety of antigens allows for
endless possibilities in application. The global market size was
valued at USD 130.9 billion in 2020, estimated to grow 223.7 billion
by the end of 2025 at a compound annual growth rate of 11.31\%\cite{market2020}. Four of the top 10 best-selling drugs in 2020 were
monoclonal antibodies\cite{Urquhart2021}.

In order to rationally design therapeutic antibodies, knowledge of
their structure is essential. The acquired structural information can
be used to increase binding affinity to a target of interest,
predicting both the exact binding site and the antibody stability as
well as assessing immunogenicity\cite{Abhinandan2007}. As
experimental structure determination is very costly and time
consuming, computational predictions of an antibody's structure are
used to streamline the process.

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly achievable\cite{North2011,Weitzner2015}. However, the CDR loop 3 of the
heavy chain (CDR-H3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable\cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops\cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-Angstrom non-antibody
structural neighbour, as well as that 30\% of CDR-H3 loops have a
completely unique structure compared with under 3\% for all other
loops on average\cite{Regep2017}.

Apart from being the most structurally diverse, the CDR-H3 loop is also
the most important for antigen binding, being located at the center of
the binding site and forming the most contacts to the antigen\cite{MacCallum1996}. In fact, it was demonstrated that differences in
this loop alone were sufficient to enable otherwise identical
antibodies to distinguish between various antigens\cite{Xu2000}.

@@

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDR-H3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDR-H3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above\protect\cite{Wong2015}.}  
  \label{fig:loopdist}
\end{figure}

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDR-H3 loops. This was also shown by the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build structural models from antibody sequences. The CDR-H3 loop
modelling quality achieved at the contests was on average much lower
for loops of longer lengths\cite{Almagro2011,Almagro2014}.

Several different approaches for generating three-dimensional
structure models from antibody sequences exist such as RosettaAntibody\cite{Sircar2009,Sivasubramanian2009},
ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017} and abYmod,
UCL's in-house software developed by Prof. Martin. One of the most
used methods is RosettaAntibody, which implements template selection
and ab initio CDR-H3 loop modelling using loop fragments and employing
specific angle restraints which bias the conformational space towards
so-called `kinked' loops\cite{Schoeder2021,Weitzner2017}. In
contrast, ABodyBuilder uses a database search algorithm
(FREAD\cite{Choi2010}) for CDR loop modelling.

In this project abYmod is used, which is to be found at
http://abymod.abysis.org/. The program utilizes extensive canonical
class definitions, VH/VL angle prediction and a large database of loop
structures (LoopDB) for CDR-H3 modelling to achieve optimal
results. Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues\cite{Martin1996}and where
an exact match is not possible, a nearest match is made. Then, the
program identifies the 10 best overall-matching PDB files according to
sequence identity for the light and heavy chain. Of these, the best
template is then identified for each CDR using sequence similarity and
identity. In general, modelling the antibody using the single best
overall-matching template works best and the CDR-specific templates
are used if there is no canonical match. The VH/VL packing angle is
then determined either using machine learning or the chosen template
structure for one of the chains. CDR specific templates, if selected,
are grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using the LoopDB database,
containing CDR-H3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results or outperforms other modelling software
(see results section).

@@

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.CDR-H2) of the antibody
model. The given score is the probability that a specific region
(e.g. CDR-H2) will be will be modeled within a specific RMSD threshold\cite{Leem2016}. Thus, it can be used to obtain an expected RMSD
value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data\cite{Leem2016}. ABLooper also provides a
confidence metric for the CDR-H3 loop model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}. However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality model. Furthermore, it remains unclear
how well the generated diversity score reflects model
quality\cite{Abanades2022}.

@@

\section{Results}
To asses the predictive power regarding the CDR-H3 loop, the software
was tested on a test-set of antibody structures used in the 2014 and
2011 Antibody Modelling Assessments
\cite{Almagro2011,Almagro2014}. As the results depicted in
\ref{fig:AMA} show, abYmod achieves results similar to, or better than
other modelling programs. However, the outliers with very high RMSD
values increase abYmod's RMSD average. The predictor in this work
would aim to identify such outlier models.

The predictive power of any machine learning model is largely
dependent on the quality and size of the dataset it was trained on. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{XXXX},
established by Prof. Andrew Martin, was utilised (1924 non-redundant
structures). The root-mean-square deviation (RMSD) value, a measure of
distance between backbone C-$\alpha$ atoms of superimposed crystal structures
and modelled structures, is calculated (see methods). This metric for
model quality was used to classify models.

@@

The full pipeline for creating the final machine learning model that
will predict model quality by giving its RMSD range starts with a
feature-set calculation using the antibody sequence. The feature set
includes many different attributes linked to sequence, physical
characteristics, interactions, etc. within as well as outside of the
loop. There is a plethora of information to train our classifier on,
including packing quality, protrusion, hydrophobicity, pseudo-energy
(see protrusion, hydrophobicity, pseudo-energy (see methods for
details). The sequence logo \ref{fig:logo} visualizes amino acid
occurrence within the loop sequence, elements of which can be
extracted as features \cite{Thomsen2012,Shaner1993}.

@@

A validation set, usually used for testing during the optimization
stage will be omitted in favour of stratified K-folds cross-validation
(k=10)\cite{Krstajic2014,Kohavi1995}.

\subsection{Model Assessment} 
Model assessment must be considered at two levels as performance
metrics of binary and multi-class classifiers are calculated
differently and must thus be considered separately. The Mathews
Correlation Coefficient (MCC)\cite{Matthews1975} is deemed the most
informative, taking the ratios of the four confusion matrix categories
into account 39and is thus more reliable than the F1 score and
accuracy . It is also consistent for both binary and multi-class
problems and therefore well suited for our purpose.

@@

\begin{figure}
  \includegraphics[scale=0.6]{angle.eps}
  \caption {Diagram visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e H95 and H102, shown as
    red spheres) of the CDR-H3 (shown in red) are identified. Then, a
    line is drawn between the two C$\alpha$ atoms of these residues (yellow
    dashed line). The distance of the C$\alpha$-atom of each residue in the
    CDR-H3 loop to this line is calculated. The residue which has the
    greatest distance to the line (shown as green sphere) is outputted
    as one-letter amino acid code and used as feature. The distance in
    \AA\ (depicted as yellow arrow) is used as the `protrusion'
    feature. }
  \label{fig:angle}
\end{figure}
